---
title: "PA1_template"
author: "Xinli"
date: "May 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Loading and preprocessing the data

```{r}
setwd("~/Desktop/reproducible_research_coursera")
dat <- read.csv("activity.csv")
dat$date <- as.Date(dat$date, format="%Y-%m-%d")
```

* What is mean total number of steps taken per day

```{r, message=FALSE, warning=FALSE}
library(dplyr)
total_steps <- dat %>% group_by(date) %>%summarize(total_steps=sum(steps, na.rm=T))
```

```{r, fig.height=4, fig.width=5}
hist(total_steps$total_steps, breaks=30,main = "Total number of steps taken per day", xlab = "total steps")
mean <- mean(total_steps$total_steps, na.rm = T)
mean
med <- median(total_steps$total_steps, na.rm = T)
med
```

So the mean of the total number of steps taken per day is `r mean`, median is `r med`. 

* What is the average daily activity patter?

```{r,fig.height=4, fig.width=5}
average_interval <- dat %>% group_by(interval) %>% summarize(average=mean(steps,na.rm=T))
plot(average_interval$interval, average_interval$average, type = 'l', xlab = "5-minute interval", ylab="average number of steps taken", main = "Time Series Plot")
maximum = which.max(average_interval$average)
average_interval[maximum, ]
```

On average across all the days in the dataset, interval `r average_interval[maximum, ][1]` contains the maximum nnumber of steps, which is `average_interval[maximum, ][2]`. 

* Imputing missing values

```{r}
total_missing <- sum(is.na(dat$steps))
total_missing
```

So the total number of missing values in the dataset is `r total_missing`.

Filling in the missing values using mean for that 5-minute interval.

```{r}
combined <- merge(dat, average_interval, by="interval", all.x = T)
combined$steps <- ifelse(is.na(combined$steps), combined$average, combined$steps)
dat.1 <- combined[,1:3]
```

```{r, fig.height=4, fig.width=5}
total_steps.1 <- dat.1 %>% group_by(date) %>% summarise(total=sum(steps))
hist(total_steps.1$total,breaks = 30, main = "Total number of steps taken per day",xlab = "total steps")
mean.1 <-mean(total_steps.1$total)
med.1 <- median(total_steps.1$total)
```

After imputing missing values using means of 5-minute intervals, the mean of total number of steps taken per day in new dataset is `r format(mean.1, digits=2)`, the median is `r format(med.1, digits=2)`. 

The impact of imputing missing data on the estimates of the total daily number of steps is the mean of total number of steps equals to the median.

* Are there differences in activity patterns between weekdays and weekends?

```{r,fig.height=4, fig.width=5}
dat.1$weekday <- weekdays(dat.1$date)
dat.1$weektype <- ifelse(dat.1$weekday %in% c("Saturday", "Sunday"), "weekend", "weekday")
average_steps <- dat.1 %>% group_by(weektype, interval) %>%
    summarize(average=mean(steps))
library(ggplot2)
p <- ggplot(average_steps, aes(x=interval,y=average,color=weektype))+
    geom_line()+
    facet_wrap(~weektype,ncol = 1, nrow = 2)
p
```


