---
title: "Fatalities, Injuries, Property Damage from Severe Weather"
author: "Xinli"
date: "June 7, 2016"
output: html_document
---

# Synopsis

Using U.S. National Oceanic and Atmospheric Administration's (NOAA) storm data to analyze fatalities, injuries and property damage caused by stroms and other severe weather events in U.S..

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Processing

Data is downloaded from [link](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)

Documentation of data is provided by [National Weather Service](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)

```{r,echo=TRUE}
setwd("~/Desktop")
datFileName <- "./repdata-data-StormData.csv.bz2"
if(!file.exists(datFileName)){
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
    download.file(url=url, destfile = datFileName)
}
dat <- read.csv("repdata-data-StormData.csv.bz2")
```

```{r, message=FALSE,echo=TRUE}
library(dplyr)
fatal <- dat %>% group_by(EVTYPE) %>% 
    summarize(fatal_total=sum(FATALITIES)) %>% arrange(desc(fatal_total))
injury <- dat %>% group_by(EVTYPE) %>%
    summarize(injury_total=sum(INJURIES)) %>% arrange(desc(injury_total))
```

```{r}
magnitude <- function(dem,exp){
    amount <- 0
    if (exp=='K'| exp=='k'){amount=dem*1000}
    else if (exp=='M'|exp=='m'){amount=dem*1000000}
    else if (exp=='B'){amount=dem*1000000000}
    return(amount)
}
dat_subset <- 
    subset(dat, PROPDMG>0|CROPDMG>0)[c("EVTYPE","PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")]

dat_subset$total_loss <- 
with(dat_subset,
     mapply(magnitude,PROPDMG,PROPDMGEXP)+mapply(magnitude,CROPDMG,CROPDMGEXP))
damage <- dat_subset %>% group_by(EVTYPE) %>%
    summarize(total_damage=sum(total_loss)) %>%
    arrange(desc(total_damage))
```

## Results

```{r, echo=TRUE, message=FALSE,fig.height=3,fig.width=6}
library(ggplot2)
p<-ggplot(as.data.frame(fatal[1:5,]), aes(x=EVTYPE,y=fatal_total))+
    geom_bar(stat="identity",fill="aquamarine1")+
    coord_cartesian(ylim = c(0, 6000))+
    geom_text(aes(label = fatal_total), vjust = -0.3, size = 3.5)+
    xlab("type of events")+
    ylab("total fatal amount")+
    ggtitle("Total Fatal by Severe Weather Events")
p
```

From barplot, tornado causes most fatal. The rest is excessive heat, flash flood, heat and lighting.

```{r, echo=TRUE, fig.height=3,fig.width=6}
p<-ggplot(as.data.frame(injury[1:5,]), aes(x=EVTYPE,y=injury_total))+
    geom_bar(stat="identity",fill="#E46726")+
    coord_cartesian(ylim = c(0, 100000))+
    geom_text(aes(label = injury_total), vjust = -0.3, size = 3.5)+
    xlab("type of events")+
    ylab("total injuries amount")+
    ggtitle("Total Injuries by Severe Weather Events")
p
```

From above barplot, tornado causes most injuries. The rest is tstm wind, flood, excessive heat and lighting.

From fatal and injury data, tornado is the most harmful with respect to population health.

```{r}
damage$total_damage <- damage$total_damage / 10^6
p<-ggplot(as.data.frame(damage[1:5,]), aes(x=EVTYPE,y=total_damage))+
    geom_bar(stat="identity",fill="steelblue")+
    coord_cartesian(ylim = c(0, 160000))+
    geom_text(aes(label = total_damage), vjust = -0.3, size = 3.5)+
    xlab("type of events")+
    ylab("total property damage (Million)")+
    ggtitle("Total Property Damage by Severe Weather Events")
p
```

From above barplot, flood causes most property damages. That is to say, flood has the greates economic consequences.
